Viscous Transport in Eroding Porous Media

### 2020-01-02 docs/transport/Response_to_Reviewer_2.tex

\end{itemize}

\noindent
\comment{p.10, end of section 4.  The authors do not explicitly say how
pore size is calculated.  Is it simply the length of the portion of a
Delaunay side that lives outside the grains?}
\begin{itemize}
 \item In Figure 3, we clarify the definition by adding the statement:
    {\em represented by the length of the red line segments}.
  \item In the text, we now state:
    \begin{quotation}
    \noindent
      Finally, we define the pore sizes to be the minimum distance
      between all pairs of neighboring grains, and these distances are
      represented by the red line segments in figure 3(a).
    \end{quotation}
\end{itemize}

\noindent
\comment{p.17 ``single body close to a wall", the 'experiment' mentioned
here is, I believe, a numerical `experiment'.  It might be good to
indicate that clearly.}
\begin{itemize}
  \item The reviewer is correct that this is a numerical experiment.

  \item There were other instances in the paper where we claimed to be
    performing an `experiment'. We have clarified each of the instances
    by referring to them as a `numerical experiment'.
\end{itemize}

@@

  \item We have updated the caption to read:
    \begin{quotation}
      \noindent
      200 particle trajectories in the second geometry from figure 6.
      The particles are initially equispaced at $(−1,y)$, where $y \in
      (−1, 1)$. The first five snapshots are equispaced in time. The
      bottom right frame is a magnification of the fifth snapshot, but
      with additional trajectories.
    \end{quotation}

  \item We have also updated the paragraph describing figure 8 to
    clarify that it is showing tracer trajectories that trace out
    streamlines.
\end{itemize}

\noindent
\comment{Figure 9.  While the sharp features in subfigure (b) are
apparently finite cusps, they raise the question of whether any infinite
dwell times (divergent $\tau$) are ever observed in these simulations -
i.e.~are particles ever trapped inside recirculating zones?}
\begin{itemize}
  \item Thank you for pointing out this possibility. The
    recirculation zones are close to the eroding bodies and are
    typically very small. Because we measure the local tortuosity by
    initiating streamlines significantly far from the grains, each of
    them connect the inlet to the outlet.

  \item We have updated the text to read:
    \begin{quotation}
      \noindent
      We note that if a streamline originates in a recirculation zone,
      which is possible in two dimensions (Higdon 1985), then the local
      tortuosity will be infinite. However, in each of our numerical
      experiments, all the streamlines connect the inlet to the outlet.
      Having only finite local tortuosities, the hydraulic tortuosity is
      defined by taking the average over all points on the inlet
      cross-section
    \end{quotation}
\end{itemize}

### 2020-01-02 docs/transport/Response_to_Reviewer_3.tex

\noindent
Dear Silas Alben,
\\ \\
\noindent
Thank you for handling the manuscript, and we would like to thank the
reviewer for their positive and constructive comments.  The attached
manuscript addresses all the reviewer's comments and the result is what
we believe is a stronger manuscript.  We also note that formatting
issues were resolved in the bibliography. An itemized list of the
changes addressing the reviewer's comments are below.
\\ \\
\noindent
Sincerely,
\\
\noindent
Shang-Huan Chiu, Nick Moore, and Bryan Quaife

\section*{Response to Reviewer 3}

\noindent
\comment{Pores in natural porous media care of three-dimensional
geometry.  The authors may include some discussion on how to extend
their numerical method to three-dimensional problems.  In addition, they
may also discuss the limitations of the erosion analysis in
2-dimensional flow to be applied for three-dimensional flow. If the
erosion process in three-dimensional flow is very different from that in
two-dimensional flow, the analysis of this paper may have less relevant
to the analysis of erosion in natural porous media.}
\begin{itemize}
  \item The reviewer is correct that by assuming the geometry is
    two-dimensional, we have limited the applicability for natural
    porous media. There are several numerical challenges to simulate
    three-dimensional erosion, with the most significant being the
    $\theta-L$ formulation, quadrature method, spectral representation
    of the geometry, and the boundary integral equation solver.

  \item In three-dimensions, others have extended the $\theta-L$
    formulation to remove stiffness in interfacial flow, spectrally
    represented geometries, and efficiently solving BIEs.  The challenge
    with these methods is their computational cost. To consider
    suspensions with the resolution and number of grains that we
    consider, the BIE would most likely require a parallel
    implementation that would run on high-performance computing
    architectures. At this point, this is not the goal of our work.
    Rather, we are interested in using very high-order methods to
    provide an analysis of the effects of erosion on transport which
    seems to be missing from the literature.

  \item The Barycentric quadrature method relies on a complex variable
    formulation, so it is inherently two-dimensional. However, there are
    other quadrature methods that can be applied in three dimensions,
    but they are also computationally expensive. For example, if the
    geometry is represented with spherical harmonics, a
    Gaussian-Legendre quadrature method requires $\mathcal{O}(p^3)$
    operations, where $p$ is the highest-order spherical harmonic.

@@

\end{itemize}

\noindent
\comment{In the numerical examples of the paper the grains are all
separated. In natural porous media, grains are compressed to form grain
contacts. The authors may include discussion on numerical challenges in
studying erosion process in natural porous media.}
\begin{itemize}
  \item The reviewer is correct that porous media grains are compressed
    and form grain contact. This is the result of processes such as
    sedimentation and transport of the grains themselves. This, again,
    is a limitation of our method since we only consider immobile
    bodies.

  \item We plan to consider mobile bodies in the future, and this will
    require a significant computational effort.

  \item To the best of our knowledge, nobody has studied the effect
    erosion has on transport through a porous media. Therefore, as a
    first step, we limit ourselves to two-dimensional suspensions of
    immobile bodies, and we plan to remove these limitations in the
    future. In the Introduction, we have added the following paragraph:
    \begin{quotation}
      \noindent
      Since high-order methods have never been used to study transport
      in eroding porous media, we consider, as a first step,
      two-dimensional and immobile grains. These assumptions allow us to
      simulate large densely-packed suspensions (up to $100$ in this
      manuscript), and to clearly interpret the results. More realistic
      three-dimensional and mobile configurations would require
      overcoming a few challenges. First, a three-dimensional extension
      of the $\theta-L$ formulation exists (Ambrose {\em et al.}~2013),
      but the Barycentric quadrature method used here relies on a
      complex variable formulation of the BIE, making it inherently two-
      dimensional. Therefore, a different quadrature method would be
      required. Additional challenges include spectral representations
      of three-dimensional geometries and efficiently solving
      discretizations of the BIE. While high-order methods for
      three-dimensional BIEs have been developed, their computational
      cost limits the number of bodies that can be practically
      simulated. Lastly, mechanisms such as sedimentation play a role in
      realistic media and affect the pore structure. Simulating mobile
      grains introduces the additional challenge of resolving contact,
      which we plan to address in future work.
    \end{quotation}
\end{itemize}
